---
title: "integration_supplementary"
author: "Ali Balubaid"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r activate conda env}
# Multiple integration tools were devekloped in python. We use reticulate with a conda environment where all relevant packages are installed.

library(reticulate)
# Seeing your enviroments
env_names = conda_list()

#Using it
# conda_list()[[1]][10] 
# env_names$%>% 
use_condaenv("/home/balubao/anaconda3/envs/blastoid_integration_env/bin/python", required = TRUE)
```

```{r}
# set-up

# renv::deactivate()
setwd('/home/balubao/Documents/Research/blastoid_comparison/')
source("script/Initiation.R", verbose = FALSE)

require(gridExtra)
require(grid)
require(lattice)
```

# Supplementary file

## Comment from Reviewer
The integration of multiple public datasets is one of the key parts in this study. 
Therefore, the customized integration methods and parameters used to mitigate batch effects should be more thoroughly explained.

In response to the reviewer's request, we further explore different integration schemes, cluster the integrated data, and analyze the distribution of cell-types across the clusters.



Top integration schemes:
1. anchor-based SNN integration (RCPA)
2. CCA alignment integration (CCA)
<!-- 3. Scanorama snn (scanorama gene) -->
3. MNN-based integratoin mnn (fastMNN)

<!-- Try the following: -->
<!-- Seurat v3 CCA -->
<!-- Seurat v3 RPCA -->
<!-- scanorama (gene) -->
<!-- ?scVI -->
<!-- scGen, scANVI, scVI -->



## Data Preprocessing

We load the preprocessed annotated data for integration.

```{r load data}
library(SeuratDisk)

print("Loading data")
seurat.list = readRDS("annotated_preintegrated_seurat_list.rds")


```

```{r}
# colnames(seurat.list$Fan2021) = gsub("-1","",sapply(strsplit(colnames(seurat.list$Fan2021), split = "_"),"[[",4))
# seurat.list$Fan2021 = AddMetaData(Fan2021, metadata = seurat.list$Fan2021@meta.data[grep("annot", colnames(seurat.list$Fan2021@meta.data) , value = TRUE)])

# colnames(seurat.list$Liu2021) = gsub("-1","",sapply(strsplit(colnames(seurat.list$Liu2021), split = "_"),"[[",4))
# seurat.list$Liu2021 = AddMetaData(Liu2021, metadata = seurat.list$Liu2021@meta.data[grep("annot", colnames(seurat.list$Liu2021@meta.data) , value = TRUE)])

# colnames(seurat.list$Yu2021_5iLA) = gsub("-1","",sapply(strsplit(colnames(seurat.list$Yu2021_5iLA), split = "_"),"[[",4))
# seurat.list$Yu2021_5iLA = AddMetaData(yu.list$Yu2021_5iLA, metadata = seurat.list$Yu2021_5iLA@meta.data[grep("annot", colnames(seurat.list$Yu2021_5iLA@meta.data) , value = TRUE)])

# colnames(seurat.list$Yu2021_PXGL) = gsub("-1","",sapply(strsplit(colnames(seurat.list$Yu2021_PXGL), split = "_"),"[[",4))
# seurat.list$Yu2021_PXGL = AddMetaData(yu.list$Yu2021_PXGL, metadata = seurat.list$Yu2021_PXGL@meta.data[grep("annot", colnames(seurat.list$Yu2021_PXGL@meta.data) , value = TRUE)])

```



```{r prepare data}

# make features the same by turning back to gene IDs

# for SS2, use feature metadata

# for 10x, retrieve feature csv file
feature.files = system(paste0("ls ../Data/*/","filtered_*_features.tsv.gz"), intern = T)
feature.tables = lapply(feature.files, read.table)
names(feature.tables) = make.unique(basename(dirname(feature.files)))

df_ds = data.frame(seurat = c("Fan2021","Liu2021","Sozen2021","Yu2021_5iLA","Yu2021_PXGL"), 
                   feat = c("Fan2021","Liu2021","Sozen2021","Yu2021","Yu2021.1"))
df_ds = df_ds[c(2,4,5),]
feature.tables = feature.tables[c(2,4,5)]

for(i in seq_along(feature.tables)){
  
  #match rownames to feature names
  feat.idx = match(rownames(seurat.list[[df_ds$seurat[i]]]), make.unique(feature.tables[[df_ds$feat[i]]]$V2))
  
  # get first two columns, gene_ID and gene name
  ftb = feature.tables[[df_ds$feat[i]]][feat.idx,seq(2)]
  
  # rename columns
  colnames(ftb) = c("ensembl_gene_id", "gene_names")
  
  # assemble transferable data.frame
  f.meta = cbind(ftb, seurat.list[[df_ds$seurat[i]]][["RNA"]][[]])
  rownames(f.meta) = rownames(seurat.list[[df_ds$seurat[i]]])
  
  # transfer
  seurat.list[[df_ds$seurat[i]]][["RNA"]][[]] = f.meta
}

# turn into assay5 objects to change rownames
for(i in seq_along(seurat.list)){
  if(class(seurat.list[[i]][["RNA"]]) != "Assay5"){
    seurat.list[[i]][["RNA"]] <- as(object = seurat.list[[i]][["RNA"]], Class = "Assay5")
  }
}

# drop features with no ENSEMBL ID
# seurat.list = lapply(seurat.list, function(x){ 
#   x = x[!is.na(x[["RNA"]][[]]$ensembl_gene_id),]
#   return(x)})

# set all feature names to gene_IDs
seurat.list = lapply(seurat.list, function(x){
  newnames = x[["RNA"]][[]]$ensembl_gene_id
  newnames[is.na(newnames)] = rownames(x[["RNA"]][[]])[is.na(newnames)]
  rownames(x) = newnames
  return(x)})

# sanity check
lapply(seurat.list, function(x) head(rownames(x)))
lapply(seurat.list, function(x) table(is.na(rownames(x))))
```


```{r}


# print("Updating data structure")
# seurat5.list = lapply(seurat.list, function(obj){
#   obj[["SCT5"]] <- as(object = obj[["SCT"]], Class = "Assay5")
#   return(obj)
# })

print("Merging data")
obj = merge(seurat.list[[1]], seurat.list[seq(2,length(seurat.list))])

print("Splitting data")
DefaultAssay(obj) = "RNA"

obj = JoinLayers(obj)

obj = split(obj, assay = "RNA", f = obj$sample)
obj@assays$RNA@layers[grep("scale.data",names(obj@assays$RNA@layers))] = NULL

print("Processing data")
plan("multicore", workers = 32)
obj <- FindVariableFeatures(obj)
obj$ERCC.perc[is.na(obj$ERCC.perc)] = 0
obj = ScaleData(obj, vars.to.regress = c("ERCC.perc","S.Score", "G2M.Score", "nFeature_RNA", "nCount_RNA"))
obj <- RunPCA(obj)
obj <- FindNeighbors(obj, dims = 1:30, reduction = "pca")
split_obj <- FindClusters(obj, resolution = 2, cluster.name = "unintegrated_clusters")
# split_obj = split_obj[rowSums(split_obj)>0,]

save(split_obj, file = "preintegrated_seurat_obj.rda")
# saveRDS(split_obj, "preintegrated_seurat_obj.rds")
```

```{r}
load("preintegrated_seurat_obj.rda")
```


## Seurat CCA

Our first integration method is the canonical correlation analysis (CCA) seurat integration method. This method identifies the CC components, which represent the dimensions with most variability, and does not require each to be perpindicular as with principal component analysis (PCA).  

```{r Seurat V3 CCA}
library(Seurat)
library(SeuratWrappers)

cca_int_obj = IntegrateLayers(object = split_obj, method = CCAIntegration, assay = "RNA",
                              orig.reduction = "pca", new.reduction = "integrated.cca", 
                              verbose = FALSE
)



```


## Seurat RPCA

Our second integration method uses reciprocal principal component analysis (RPCA). 

```{r Seurat V3 RPCA}

setwd('/home/balubao/Documents/Research/blastoid_comparison/')

library(Seurat)
library(SeuratWrappers)

load("preintegrated_seurat_obj.rda")

rpca_int_obj <- IntegrateLayers(
  object = split_obj, method = RPCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.rpca",
  normalization.method = "LogNormalize",
  k.weight = 80,
  verbose = TRUE
)

rpca_ref_int_obj <- IntegrateLayers(
  object = split_obj, method = RPCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.rpca",
  normalization.method = "LogNormalize", reference = seq(3),
  k.weight = 80,
  verbose = TRUE
)

save(rpca_int_obj, rpca_ref_int_obj, file = "rpca_integrated_blastoid.rda")


```

## QC Integrated Data
```{r}
# find clusters
obj <- FindNeighbors(obj, reduction = "integrated.cca", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "cca_clusters")

obj <- FindNeighbors(obj, reduction = "integrated.rpca", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "rpca_clusters")

obj <- RunUMAP(obj, reduction = "integrated.cca", dims = 1:30, reduction.name = "umap.cca")
p1 <- DimPlot(
  obj,
  reduction = "umap.cca",
  group.by = c("Method", "predicted.celltype.l2", "cca_clusters"),
  combine = FALSE, label.size = 2
)

obj <- FindNeighbors(obj, reduction = "integrated.scvi", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "scvi_clusters")

```

```{r MNN}


mnn_int_obj <- IntegrateLayers(
  object = split_obj, method = FastMNNIntegration,
  new.reduction = "integrated.mnn",
  verbose = TRUE
)
saveRDS(mnn_int_obj, "mnn_integrated_blastoid.rds")


```

```{r Harmony}

harmony_int_obj <- IntegrateLayers(
  object = split_obj, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "integrated.harmony",
  normalization.method = "LogNormalize",
  k.weight = 80,
  verbose = TRUE
)
saveRDS(harmony_int_obj, "harmony_integrated_blastoid.rds")
```

```{ LIGER}
library(Seurat)
library(SeuratWrappers)
library(rliger)

pbmcsca <- ScaleData(pbmcsca, split.by = "Method", do.center = FALSE)
pbmcsca <- RunOptimizeALS(pbmcsca, k = 20, lambda = 5, split.by = "Method")
pbmcsca <- RunQuantileNorm(pbmcsca, split.by = "Method")
# You can optionally perform Louvain clustering (`FindNeighbors` and `FindClusters`) after
# `RunQuantileNorm` according to your needs
pbmcsca <- FindNeighbors(pbmcsca, reduction = "iNMF", dims = 1:20)
pbmcsca <- FindClusters(pbmcsca, resolution = 0.3)
# Dimensional reduction and plotting
pbmcsca <- RunUMAP(pbmcsca, dims = 1:ncol(pbmcsca[["iNMF"]]), reduction = "iNMF")
DimPlot(pbmcsca, group.by = c("Method", "ident", "CellType"), ncol = 3)

```

```{ scanorama}
# !pip install scvi
library(reticulate)
library(tidyverse)

# 
scanorama = import("scanorama")

# Integration.
counts.ls = lapply(seurat.list, function(seurat.obj){
counts = as.matrix(seurat.obj[["RNA"]]$data)
return(t(counts))
})
genes.ls = lapply(counts.ls, colnames)
counts.ls = lapply(counts.ls, as.matrix)

names(counts.ls) = NULL

# alignmnet.out = scanorama$find_alignments(counts.ls, knn = 20L)
integrated.data <- scanorama$integrate(counts.ls, genes.ls)

print("Done integrating")

```

```{ scVI}
# !pip install scvi
library(reticulate)
library(tidyverse)
library(SeuratWrappers)
#Checking python
# scvi = import("scvi")
# scib = import("scib")

scvi_int_obj <- IntegrateLayers(
object = split_obj, 
method = scVIIntegration,
orig.reduction = "pca",
new.reduction = "integrated.scvi",
conda_env = "/home/balubao/anaconda3/envs/blastoid_integration_env", verbose = TRUE
)
saveRDS(scvi_int_obj, "scvi_integrated_blastoid.rds")

```


## Evaluate Integration
<!-- https://www.nature.com/articles/s41592-021-01336-8#Sec11 -->
To evaluate the different integration methods, we use the metrics introduced in scib listed below:

### Biological Conservation
* Cell type ASW
* Cell cycle conservation
* Graph cLISI
* Adjusted rand index (ARI) for cell label
* Normalised mutual information (NMI) for cell label
* Highly variable gene conservation
* Isolated label ASW
* Isolated label F1
* Trajectory conservation

### Batch Correction
* Batch ASW
* Principal component regression
* Graph iLISI
* Graph connectivity
* kBET (K-nearest neighbour batch effect)

```{r scVI}
# !pip install scvi
library(reticulate)
library(tidyverse)

#Checking python
# scvi = import("scvi")
scib = import("scib")

scib.metrics
```



```{python}
# !pip install scib
# !pip install 'scib[rpy2,bbknn]'

import scib

scib.metrics.


```



```{r}
seurat.list = readRDS("annotated_seurat_list.rds")
Integrated.seurat = LoadH5Seurat("intregrated_all_4.h5seurat")

## transfer annotation labels
annotation_labels = sapply(seurat.list, function(x) x$annot_major)
annotation_labels = unlist(annotation_labels, use.names = TRUE)
names(annotation_labels) = sub("^[^.]*\\.", "", names(annotation_labels))
Integrated.seurat$annot_major = annotation_labels[match(names(annotation_labels), colnames(Integrated.seurat))]

# generate plotting.data
plotting.data = cbind(Integrated.seurat@meta.data, Embeddings(Integrated.seurat, reduction = 'umap'), t(GetAssayData(Integrated.seurat[["SCT"]])[unlist(AllMarkers),]))
plotting.data$cluster = plotting.data$integrated_snn_res.2

df_ds = data.frame(
  # sample=names(seurat.list),
  sample=sapply(seurat.list, function(x){unique(x$sample)}), 
  ncells=sapply(seurat.list, ncol), 
  nfeatures = sapply(seurat.list, function(x){sum(rowSums(x@assays$RNA)>10)}),
  tech=sapply(seurat.list, function(x){unique(x$tech)}),
  ref=sapply(seurat.list, function(x){unique(x$ref)}))

```

```{r}
library(ggbreak)

plotting.data$sample_celltype = paste0(plotting.data$sample, "_", plotting.data$annot_major)
ggplot(plotting.data, aes(x = sample, y=log10(nCount_RNA)))+
  # geom_violin(aes(fill=annot_major))+
  geom_boxplot(aes(fill=annot_major))+
  coord_flip() + scale_y_break(c(5.5, 10))

library(ggridges)
ggplot(plotting.data, aes(x = log10(nCount_TPM), y = sample))+
  geom_density_ridges()


```

# Generate Figure 1 using new data.
```{r}
fig_1A_v2 = function(){
  
  Integrated.seurat@meta.data %>% group_by(sample) %>% count(annot_major) %>% mutate(n=n/sum(n))-> mat_ct_cnts
  mat_ct_cnts = reshape2::acast(mat_ct_cnts, formula = sample ~ annot_major)
  hclust_ct_cnts = hclust(dist(mat_ct_cnts))
  
  df_ds$sample = factor(df_ds$sample, levels = rownames(mat_ct_cnts)[hclust_ct_cnts$order])
  
  df_ds$`seq. platform` = df_ds$tech
  p1 = ggplot(df_ds, aes(y=sample, x=log10(ncells), fill=`seq. platform`))+
    geom_col()+
    theme_bw()+
    ylab("Dataset")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  p2 = ggplot(df_ds, aes(y=sample, x=nfeatures/10000, fill=`seq. platform`))+
    geom_col()+
    theme_bw()+
    xlab("nfeatures (10k)")+
    # theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
  # p_sctype = ggplot(Integrated.seurat@meta.data, aes(x=customclassif, fill=sample))+
  # geom_bar(position = "fill")+
  # theme_bw()+
  # theme(axis.text.x = element_text(angle=45, vjust = 1, hjust = 1))
  
  tb1 = table(plotting.data$sample, plotting.data$annot_major)
  tb1 = sweep(tb1, 1, STATS = rowSums(tb1), "/")
  tb1 = as.data.frame.matrix(tb1)
  tb1$samples = rownames(tb1)
  tb1 = as.data.frame(tb1, row.names = rownames(tb1))
  
  pretmp = df_ds %>% mutate(log10_ncells = log10(ncells)) %>% mutate(nfeatures_10k = nfeatures/1e4)
  tmp = cbind(pretmp,tb1[match(pretmp$sample,tb1$sample),])
  
  tmp = reshape2::melt(tmp)
  tmp$sample = factor(tmp$sample, levels = rownames(mat_ct_cnts)[hclust_ct_cnts$order])
  tmp$variable = factor(tmp$variable)
  tmp$`frac. of cells` = tmp$value
  #, levels = c("ncells","nfeatures","log10_ncells","nfeatures_10k","Inner cell mass","Epiblast cells","Primitive Endoderm cells","Trophoblast cells",
  #          "Early Amnion","Late Amnion","extravillous cytotrophoblasts","Syncytiotrophoblasts","cytotrophoblasts","Unknown"))
  # tmp$sample = fct_inorder(tmp$sample)
  # p1 = tmp %>% filter(variable %in% "log10_ncells") %>% ggplot()+
  #   geom_tile(aes(y=sample, x=variable, fill=value))+
  #   scale_fill_gradient(low = "black", high = "green")+ 
  #   xlab("")+
  #   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  # p2 = tmp %>% filter(variable %in% "nfeatures_10k") %>% ggplot()+
  #   geom_tile(aes(y=sample, x=variable, fill=value))+
  #   scale_fill_gradient(low = "black", high = "purple")+ 
  #   xlab("")+
  #   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  #         axis.text.y = element_blank(),
  #         axis.ticks.y = element_blank(),
  #         axis.title.y = element_blank())
  p3 = tmp %>% filter(variable %in% levels(tmp$variable)[-c(1,2,3,4)]) %>% ggplot()+
    geom_tile(aes(y=sample, x=variable, fill=`frac. of cells`))+
    scale_fill_viridis_c()+
    # scale_fill_gradient2(midpoint=0.5, 
    # low="blue", 
    # mid="white",
    # high="red", 
    # space ="Lab")+
    xlab("") + 
    coord_fixed() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
  p4 = ggdendrogram(hclust_ct_cnts, rotate = TRUE, size = 3, labels = FALSE) + theme_void()
  
  wrap_plots(list(p1,p2,p3,p4), design = "ABCD", guides = "collect")
}
fig_1A_v2()
# ggsave("fig1A_data_summary.svg", width = 12, height = 6, dpi = "retina")
```


```{r}
fig_1B = function(){
  plotting.ls = lapply(seurat.list, function(x){tmp_plotting.data = cbind(x@meta.data, Embeddings(x, "umap"))})
  # tmp_plotting = do.call("rbind", plotting.ls)
  tmp_plotting = rbindlist(plotting.ls, use.names = TRUE, fill = TRUE)
  tmp_plotting$`annotations (auto)` = tmp_plotting$annot_major
  sampled_data <- tmp_plotting %>%
    group_by(sample) %>%
    sample_n(size = 5000, replace = TRUE) %>%
    ungroup() %>%
    distinct()  # Remove duplicates
  ggplot(sampled_data, aes(x=umap_1, y=umap_2, color=`annotations (auto)`))+
    geom_point(size=0.5)+
    theme_bw()+
    coord_fixed()+
    facet_wrap(~sample, nrow = 2)+
    theme(strip.text = element_text(size=6))
  
}
fig_1B()
ggsave("fig1B_data_UMAP.svg", width = 12, height = 4.5, dpi = "retina")
```

```{r}
p1 = DimPlot(Integrated.seurat, group.by = "seurat_clusters", label = TRUE) +     
  # theme_void() +
  theme(legend.position = "none") + 
  coord_fixed()
p2 = DimPlot(Integrated.seurat, group.by = "annot_major") + 
  theme(legend.position = "none") + 
  # scale_fill_viridis_d(option = "D")+
  coord_fixed()
p3 = DimPlot(Integrated.seurat, group.by = "sample") + 
  theme(legend.position = "none") + 
  coord_fixed()
cowplot::plot_grid(p1,p2,rel_heights=c(1,1), ncol = 1)
```

# Celltype Distributions
```{r}
plotting.data$cluster = plotting.data$rpca_clusters 
tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$annot_major))
clust_annot_hclust = hclust(dist(tmp_mat))
tmp = plotting.data
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])

p1 = ggplot(tmp, aes(x=cluster, fill=annot_major))+
  geom_bar()+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank())

p2 = ggdendrogram(clust_annot_hclust, labels = FALSE, rotate = TRUE) + 
  theme_void() + 
  scale_y_discrete(expand=c(0,0))+
  scale_y_reverse()

# p2 + p1 + plot_layout(ncol = 2, widths = c(1, 2))
cowplot::plot_grid(p2,p1,rel_widths=c(.4,1))



tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$annot_major))
clust_annot_hclust = hclust(dist(sweep(tmp_mat, 1, rowSums(tmp_mat), FUN = "/")))
tmp = plotting.data
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])


p1 = ggplot(tmp, aes(x=cluster, fill=annot_major))+
  geom_bar(position = "fill")+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  scale_y_reverse()+
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank())

p2 = ggdendrogram(clust_annot_hclust, labels = FALSE, rotate = TRUE) + 
  theme_void() + 
  scale_y_discrete(expand=c(0,0))+
  scale_y_reverse()

# p2 + p1 + plot_layout(ncol = 2, widths = c(1, 2))
cowplot::plot_grid(p2,p1,rel_widths=c(.4,1))
```

```{r}
tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$annot_major))

clust_annot_hclust = hclust(dist(tmp_mat))

sample_names <- colnames(tmp_mat)

# Obtain the dendrogram
dend <- as.dendrogram(clust_annot_hclust)
dend_data <- dendro_data(dend)

# Setup the data, so that the layout is inverted (this is more 
# "clear" than simply using coord_flip())
segment_data <- with(
  segment(dend_data), 
  data.frame(x = y, y = x, xend = yend, yend = xend))

# Use the dendrogram label data to position the gene labels
gene_pos_table <- with(
  dend_data$labels, 
  data.frame(y_center = x, cluster = as.character(label), height = 1))

# Table to position the samples
sample_pos_table <- data.frame(celltype = sample_names) %>%
  mutate(x_center = (1:n()), 
         width = 1)

# Neglecting the gap parameters
heatmap_data <- tmp_mat %>% 
  reshape2::melt(value.name = "n", varnames = c("cluster", "celltype")) %>%
  mutate(cluster = as.character(cluster)) %>%
  left_join(gene_pos_table) %>%
  left_join(sample_pos_table)

# Limits for the vertical axes
gene_axis_limits <- with(
  gene_pos_table, 
  c(min(y_center - 0.5 * height), max(y_center + 0.5 * height))
) + 
  0.1 * c(-1, 1) # extra spacing: 0.1

# Heatmap plot
plt_hmap <- ggplot(heatmap_data, 
                   aes(x = x_center, y = y_center, fill = n, 
                       height = height, width = width)) + 
  geom_tile() +
  scale_fill_gradient2("expr", high = "darkred", low = "darkblue") +
  scale_x_continuous(breaks = sample_pos_table$x_center, 
                     labels = sample_pos_table$celltype, 
                     expand = c(0, 0)) + 
  # For the y axis, alternatively set the labels as: gene_position_table$gene
  scale_y_continuous(breaks = gene_pos_table[, "y_center"], 
                     labels = rep("", nrow(gene_pos_table)),
                     limits = gene_axis_limits, 
                     expand = c(0, 0)) + 
  labs(x = "Celltype", y = "") +
  theme_bw() +
  theme(axis.text.x = element_text(size = rel(1), hjust = 1, angle = 45), 
        # margin: top, right, bottom, and left
        plot.margin = unit(c(1, 0.2, 0.2, -0.7), "cm"), 
        panel.grid.minor = element_blank())

plt_dendr <- ggplot(segment_data) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + 
  scale_x_reverse(expand = c(0, 0.5)) + 
  scale_y_continuous(breaks = gene_pos_table$y_center, 
                     labels = gene_pos_table$cluster, 
                     limits = gene_axis_limits, 
                     expand = c(0, 0)) + 
  labs(x = "Distance", y = "", colour = "", size = "") +
  theme_bw() + 
  theme(panel.grid.minor = element_blank())

library(cowplot)
plot_grid(plt_dendr, plt_hmap, align = 'h', rel_widths = c(1, 1))
```

```{r}
library(tidyverse)
library(ggdendro)
library(vegan)
library(colorspace)
library(cowplot)

tb1=t(reshape2::acast(as.data.frame(table(plotting.data$cluster, plotting.data$annot_major)), formula = Var2~Var1))

hc=hclust(dist(tb1),method="ward.D2")
hc=reorder(hc,wts=-as.matrix(tb1)%*%seq(ncol(tb1))^2) # vegan::reorder.hclust
tree=ggdendro::dendro_data(as.dendrogram(hc),type="rectangle")

p1=ggplot(ggdendro::segment(tree))+
  geom_segment(aes(x=y,y=x,xend=yend,yend=xend),lineend="round",size=.4)+
  scale_x_continuous(expand=expansion(add=c(0,.01)))+ # don't crop half of line between top-level nodes
  scale_y_continuous(limits=.5+c(0,nrow(tb1)),expand=c(0,0))+
  theme(
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.ticks.length=unit(0,"pt"), # remove extra space occupied by ticks
    axis.title=element_blank(),
    panel.background=element_rect(fill="white"),
    panel.grid=element_blank(),
    plot.margin=margin(5,5,5,0)
  )

tb1=tb1[hc$labels[hc$order],]
t2=data.frame(V1=rownames(tb1)[row(tb1)],V2=colnames(tb1)[col(tb1)],V3=as.vector(tb1))
lab=round(100*t2$V3)
lab[lab==0]=""

p2=ggplot(t2,aes(x=factor(V1,level=rownames(tb1)),y=V3,fill=V2))+
  geom_bar(stat="identity",width=1,position=position_fill(reverse=T))+
  # geom_text(aes(label=lab),position=position_stack(vjust=.5,reverse=T),size=3.5)+
  coord_flip()+
  scale_x_discrete(expand=c(0,0))+
  scale_y_discrete(expand=c(0,0))+
  # scale_fill_manual(values=colorspace::hex(HSV(head(seq(0,360,length.out=ncol(tb1)+1),-1),.5,1)))+
  theme(
    axis.text=element_text(color="black",size=11),
    axis.text.x=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    legend.position="none",
    plot.margin=margin(5,0,5,5)
  )

cowplot::plot_grid(p2,p1,rel_widths=c(1,.4))
# ggsave("a.png",height=.25*nrow(t),width=7)
```


# Dataset Distributions
```{r}
tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$sample))
clust_annot_hclust = hclust(dist(tmp_mat))
tmp = plotting.data
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])

p1 = ggplot(tmp, aes(x=cluster, fill=sample))+
  geom_bar()+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank())

p2 = ggdendrogram(clust_annot_hclust, labels = FALSE, rotate = TRUE) + 
  theme_void() + 
  scale_y_discrete(expand=c(0,0))+
  scale_y_reverse()

# p2 + p1 + plot_layout(ncol = 2, widths = c(1, 2))
cowplot::plot_grid(p2,p1,rel_widths=c(.4,1))



tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$sample))
clust_annot_hclust = hclust(dist(sweep(tmp_mat, 1, rowSums(tmp_mat), FUN = "/")))
tmp = plotting.data
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])


p3 = ggplot(tmp, aes(x=cluster, fill=sample))+
  geom_bar(position = "fill")+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  scale_y_reverse()+
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank())

p4 = ggdendrogram(clust_annot_hclust, labels = FALSE, rotate = TRUE) + 
  theme_void() + 
  scale_y_discrete(expand=c(0,0))+
  scale_y_reverse()

# p2 + p1 + plot_layout(ncol = 2, widths = c(1, 2))
cowplot::plot_grid(p4,p3,rel_widths=c(.4,1))
```

```{r}
tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$sample))
clust_annot_hclust = hclust(dist(sweep(tmp_mat, 1, rowSums(tmp_mat), FUN = "/")))
samp_annot_hclust = hclust(dist(sweep(tmp_mat, 2, colSums(tmp_mat), FUN = "/")))

tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$sample))
tmp_mat = sweep(tmp_mat, 2, colSums(tmp_mat), FUN = "/")
tmp = as.data.frame(tmp_mat)
colnames(tmp) = c("cluster", "sample", "proportion")
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])
tmp$sample = factor(tmp$sample, levels = levels(tmp$sample)[samp_annot_hclust$order])

# Now, plot with geom_col using the calculated proportions
p3 <- ggplot(tmp, aes(x=cluster, y=proportion, fill=sample)) +
  geom_col()+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        # axis.text.y = element_blank(), # Changed from axis.text.x to axis.text.y for flipped coordinates
        axis.title.x = element_blank())+
  # strip.text.x = element_text(size=5))+
  facet_grid(~sample)

# Print the plot
p3


```


```{r}

tmp_mat = as.matrix(table(plotting.data$cluster, plotting.data$annot_major))
clust_annot_hclust = hclust(dist(sweep(tmp_mat, 1, rowSums(tmp_mat), FUN = "/")))
tmp = plotting.data
tmp$cluster = factor(tmp$cluster, levels = levels(tmp$cluster)[clust_annot_hclust$order])

tmp$CellType = tmp$annot_major
tmp$DataSet = tmp$sample
tmp$Clusters = tmp$cluster

p1 = ggplot(tmp, aes(x=Clusters, fill=CellType))+
  geom_bar(position = "fill")+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  # coord_fixed(ratio = 10)+
  # scale_fill_viridis_d(option = "D")+
  scale_y_reverse()+
  theme(axis.title.y = element_text(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(), 
        legend.position = "right") 

# clust_annot_hclust = hclust(dist(sweep(tmp_mat, 1, rowSums(tmp_mat), FUN = "/")))
# tmp = plotting.data
# tmp$cluster = reorder(tmp$cluster, new.order = levels(tmp$cluster)[clust_annot_hclust$order])


p3 = ggplot(tmp, aes(x=Clusters, fill=DataSet))+
  geom_bar(position = "fill")+
  scale_y_discrete(expand=c(0,0))+
  coord_flip()+
  # coord_fixed(ratio = 10)+
  scale_fill_viridis_d(option = "H")+
  scale_y_reverse()+
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "right")


# cowplot::plot_grid(p1,p3,rel_widths=c(1,1), nrow=1)
wrap_plots(list(p1, p3), guides = "collect")
```


```{r}
# install.packages("remotes")
# remotes::install_github("davidsjoberg/ggsankey")
library(ggsankey)

# df_long = plotting.data %>%
#   make_long(cluster, sample, annot_major)
# 
# ggplot(df_long, aes(x = x, 
#                next_x = next_x, 
#                node = node, 
#                next_node = next_node,
#                fill = factor(node),
#                label = node)) +
#   geom_sankey() +  
#   geom_sankey_label(size = 3.5, color = 1, fill = "white") +
#   scale_fill_viridis_d(option = "H") +
#   theme_sankey(base_size = 16) +
#   theme(legend.position = "none")


df_long = plotting.data %>%
  make_long(sample, annot_major)

ggplot(df_long, aes(x = x, 
                    next_x = next_x, 
                    node = node, 
                    next_node = next_node,
                    fill = factor(node),
                    label = node)) +
  geom_sankey(flow.alpha = 0.5, node.color = 1) +
  geom_sankey_label(size = 3.5, color = 1, fill = "white") +
  scale_fill_viridis_d(option = "D") +
  theme_sankey(base_size = 16) +
  theme(legend.position = "none")
```

```{r}
library(ggalluvial)

tmp_df = plotting.data %>%
  group_by(cluster, annot_major, sample) %>%
  summarise(n=n())

ggplot(data = tmp_df,
       aes(axis1 = cluster,   # First variable on the X-axis
           axis2 = annot_major, # Second variable on the X-axis
           axis3 = sample,   # Third variable on the X-axis
           y = n)) +
  geom_alluvium(aes(fill = annot_major)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void()




tmp_df = plotting.data %>%
  group_by(sample, cluster, annot_major) %>%
  summarise(n=n()) %>% 
  ungroup(annot_major) %>%
  mutate(prop = n/sum(n))

ggplot(data = tmp_df,
       aes(axis1 = cluster,   # First variable on the X-axis
           axis2 = annot_major, # Second variable on the X-axis
           axis3 = sample,   # Third variable on the X-axis
           y = prop)) +
  geom_alluvium(aes(fill = annot_major), curve_type = "sigmoid") +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void()


tmp_df = plotting.data %>%
  group_by(sample, cluster, annot_major) %>%
  summarise(n=n()) %>% 
  ungroup(sample) %>%
  mutate(prop = n/sum(n))

ggplot(data = tmp_df,
       aes(axis1 = cluster,   # First variable on the X-axis
           axis2 = annot_major, # Second variable on the X-axis
           axis3 = sample,   # Third variable on the X-axis
           y = prop)) +
  geom_alluvium(aes(fill = annot_major), curve_type = "sigmoid") +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void()
```

```{}
library(SeuratDisk)

print("Loading data")
seurat.list = readRDS("annotated_seurat_list.rds")

print("Updating data structure")
seurat5.list = lapply(seurat.list, function(obj){
obj[["SCT5"]] <- as(object = obj[["SCT"]], Class = "Assay5")
return(obj)
})

print("Merging data")
obj = merge(seurat5.list[[1]], seurat5.list[seq(2,length(seurat5.list))])

print("Splitting data")
DefaultAssay(obj) = "SCT5"

obj = JoinLayers(obj)

obj[["SCT5"]] = split(obj, assay = "SCT5", f = obj$sample)

print("Processing data")
obj <- FindVariableFeatures(obj)
obj <- RunPCA(obj)
obj <- FindNeighbors(obj, dims = 1:30, reduction = "pca")
split_obj <- FindClusters(obj, resolution = 2, cluster.name = "unintegrated_clusters")

split_obj = split_obj[rowSums(split_obj)>0,]
# seurat.list = readRDS("annotated_seurat_list.rds")
# obj = merge(seurat.list[[1]], seurat.list[-1])
# obj = CreateSeuratObject(counts = )

```